---
title: "Figure 4 and Supplementary Figure 5 Generation"
output: html_notebook
---

#Gene annotation and expression matrix data: 

```{r}

#loads the gene annotation files used in this study

#genes_anno - all genes considered and their position, gene type, locus etc - Data associated with Supplementary Table 2
#protein,non_coding - subset of genes_anno based on gene_type
#all_genes_cancer,exp_cutoff_genes - gene annotation for genes that pass filtering threshold (at least 1 FPKM in 20% of samples)

load("../Data/TARGET/PanCancer_lncRNA_AnnotationFiles_030121.rda")


#loads the expression matrix (fpkm) for all samples used in this study 

#fpkm - total expression matrix of all samples and alll genes 
#fpkm_list - list of dataframes with each cancer gene exp matrix after filtering out low exp genes

load("../Data/TARGET/PanCancer_lncRNA_ExpFiles_030121.rda")

```

#Load all the lncMod results. See the lncMod_Analysis folder to see README/scripts on how these files were generated. 

```{r}

# Load all the lncmod results - loads a data frame per cancer with correlation information per lncMod triplet. 

load("~/Dropbox/ApexaFiles/PanTarget_Paper/Figure2_PanTarget_lncMod/Data/NBL_lncMod_Triplets_Combined_Permuted_0420.rda")

nbl_lncmod = all.cor
nbl_lncmod$combo = paste(nbl_lncmod$combo,nbl_lncmod$lncRNA,sep="+")
nbl_lncmod$cancer = "NBL"
nbl_lncmod$lncRNA_name = genes_anno[as.character(nbl_lncmod$lncRNA),4]
nbl_lncmod$TF_name = genes_anno[as.character(nbl_lncmod$TF),4]
nbl_lncmod$TG_name = genes_anno[as.character(nbl_lncmod$TargetGene),4]
nbl_lncmod$lnc_TF = paste(nbl_lncmod$lncRNA_name,nbl_lncmod$TF_name,sep="_")
nbl_lncmod$lnc_TG = paste(nbl_lncmod$lncRNA_name,nbl_lncmod$TG_name,sep="_")

load("~/Dropbox/ApexaFiles/PanTarget_Paper/Figure2_PanTarget_lncMod/Data/AML_lncMod_Triplets_Combined_Permuted_0420.rda")
aml_lncmod = all.cor
aml_lncmod$combo = paste(aml_lncmod$combo,aml_lncmod$lncRNA,sep="+")
aml_lncmod$cancer = "AML"
aml_lncmod$lncRNA_name = genes_anno[aml_lncmod$lncRNA,4]
aml_lncmod$lncRNA_name = genes_anno[as.character(aml_lncmod$lncRNA),4]
aml_lncmod$TF_name = genes_anno[as.character(aml_lncmod$TF),4]
aml_lncmod$TG_name = genes_anno[as.character(aml_lncmod$TargetGene),4]
aml_lncmod$lnc_TF = paste(aml_lncmod$lncRNA_name,aml_lncmod$TF_name,sep="_")
aml_lncmod$lnc_TG = paste(aml_lncmod$lncRNA_name,aml_lncmod$TG_name,sep="_")

#aml_iqr = apply(AML,1,function(x) log2(IQR(x)))
#aml_iqr[aml_iqr%in%aml_lncmod$TF]

load("~/Dropbox/ApexaFiles/PanTarget_Paper/Figure2_PanTarget_lncMod/Data/TALL_lncMod_Triplets_Combined_Permuted_0420.rda")
tall_lncmod = all.cor
tall_lncmod$combo = paste(tall_lncmod$combo,tall_lncmod$lncRNA,sep="+")
tall_lncmod$cancer = "TALL"
tall_lncmod$lncRNA_name = genes_anno[tall_lncmod$lncRNA,4]
tall_lncmod$lncRNA_name = genes_anno[as.character(tall_lncmod$lncRNA),4]
tall_lncmod$TF_name = genes_anno[as.character(tall_lncmod$TF),4]
tall_lncmod$TG_name = genes_anno[as.character(tall_lncmod$TargetGene),4]
tall_lncmod$lnc_TF = paste(tall_lncmod$lncRNA_name,tall_lncmod$TF_name,sep="_")
tall_lncmod$lnc_TG = paste(tall_lncmod$lncRNA_name,tall_lncmod$TG_name,sep="_")

load("~/Dropbox/ApexaFiles/PanTarget_Paper/Figure2_PanTarget_lncMod/Data/BALL_lncMod_Triplets_Combined_Permuted_0420.rda")
ball_lncmod = all.cor
ball_lncmod$combo = paste(ball_lncmod$combo,ball_lncmod$lncRNA,sep="+")
ball_lncmod$cancer = "BALL"
ball_lncmod$lncRNA_name = genes_anno[ball_lncmod$lncRNA,4]
ball_lncmod$lncRNA_name = genes_anno[as.character(ball_lncmod$lncRNA),4]
ball_lncmod$TF_name = genes_anno[as.character(ball_lncmod$TF),4]
ball_lncmod$TG_name = genes_anno[as.character(ball_lncmod$TargetGene),4]
ball_lncmod$lnc_TF = paste(ball_lncmod$lncRNA_name,ball_lncmod$TF_name,sep="_")
ball_lncmod$lnc_TG = paste(ball_lncmod$lncRNA_name,ball_lncmod$TG_name,sep="_")

load("~/Dropbox/ApexaFiles/PanTarget_Paper/Figure2_PanTarget_lncMod/Data/WT_lncMod_Triplets_Combined_Permuted_0420.rda")
wt_lncmod = all.cor
wt_lncmod$combo = paste(wt_lncmod$combo,wt_lncmod$lncRNA,sep="+")
wt_lncmod$cancer = "WT"
wt_lncmod$lncRNA_name = genes_anno[wt_lncmod$lncRNA,4]
wt_lncmod$lncRNA_name = genes_anno[as.character(wt_lncmod$lncRNA),4]
wt_lncmod$TF_name = genes_anno[as.character(wt_lncmod$TF),4]
wt_lncmod$TG_name = genes_anno[as.character(wt_lncmod$TargetGene),4]
wt_lncmod$lnc_TF = paste(wt_lncmod$lncRNA_name,wt_lncmod$TF_name,sep="_")
wt_lncmod$lnc_TG = paste(wt_lncmod$lncRNA_name,wt_lncmod$TG_name,sep="_")

load("~/Dropbox/ApexaFiles/PanTarget_Paper/Figure2_PanTarget_lncMod/Data/RT_lncMod_Triplets_Combined_Permuted_0420.rda")
rt_lncmod = all.cor
rt_lncmod$combo = paste(rt_lncmod$combo,rt_lncmod$lncRNA,sep="+")
rt_lncmod$cancer = "RT"
rt_lncmod$lncRNA_name = genes_anno[rt_lncmod$lncRNA,4]
rt_lncmod$lncRNA_name = genes_anno[as.character(rt_lncmod$lncRNA),4]
rt_lncmod$TF_name = genes_anno[as.character(rt_lncmod$TF),4]
rt_lncmod$TG_name = genes_anno[as.character(rt_lncmod$TargetGene),4]
rt_lncmod$lnc_TF = paste(rt_lncmod$lncRNA_name,rt_lncmod$TF_name,sep="_")
rt_lncmod$lnc_TG = paste(rt_lncmod$lncRNA_name,rt_lncmod$TG_name,sep="_")

pantarget = list(AML,TALL,BALL,RT,WT,NBL)
names(pantarget) = c("AML","TALL","BALL","RT","WT","NBL")
pantarget_lncmod = rbind(aml_lncmod,tall_lncmod,ball_lncmod,nbl_lncmod,wt_lncmod,rt_lncmod)

#annotate direction of change for each of the lncRNA modulators

pantarget_lncmod$lncMod_type = "None"

pantarget_lncmod[abs(pantarget_lncmod$Bottom_cor)<abs(pantarget_lncmod$Top_cor),]$lncMod_type = "Enhance"
pantarget_lncmod[abs(pantarget_lncmod$Bottom_cor)>abs(pantarget_lncmod$Top_cor),]$lncMod_type = "Attenuate"

pantarget_lncmod[(pantarget_lncmod$Top_cor>0.4 & pantarget_lncmod$Bottom_cor<(-0.4)) | (pantarget_lncmod$Top_cor<(-0.4) & pantarget_lncmod$Bottom_cor>0.4),]$lncMod_type = "Inverts"

pantarget_lncmod$lnc_chr = genes_anno[as.character(pantarget_lncmod$lncRNA),5]
pantarget_lncmod$TG_chr = genes_anno[as.character(pantarget_lncmod$TargetGene),5]

#overview of lncRNA and target gene being on the same chromosome or not - add to a table

#Supplementary Table 11:

#save(pantarget_lncmod,file="PanTarget_lncMod_Results_0420.rda")

```


# Figure 4b - the types of lncRNA modulators that there are across cancers 

```{r}
#CHUNK4 - Plot per cancer the number of lncRNA modulators that either enhancer, attenuate, or invert TF-target gene correlation. The cutoffs were taken from the lncMap paper. 

#show the lncModulator type as a ratio of total number of triplets per cancer 

pantarget_lncmod_type = as.data.frame.matrix(table(pantarget_lncmod$cancer,pantarget_lncmod$lncMod_type))
pantarget_lncmod_type$total = rowSums(pantarget_lncmod_type)
pantarget_lncmod_type$Attenuate = pantarget_lncmod_type$Attenuate/pantarget_lncmod_type$total
pantarget_lncmod_type$Enhance = pantarget_lncmod_type$Enhance/pantarget_lncmod_type$total
pantarget_lncmod_type$Inverts = pantarget_lncmod_type$Inverts/pantarget_lncmod_type$total
pantarget_lncmod_type$total = NULL
pantarget_lncmod_type$cancer = rownames(pantarget_lncmod_type)
pantarget_lncmod_type = melt(pantarget_lncmod_type)
  
pantarget_lncmod_type$cancer = factor(pantarget_lncmod_type$cancer,levels=c("AML","BALL","TALL","NBL","WT","RT"))


yy = ggplot(pantarget_lncmod_type,aes(cancer,value,fill=variable))+geom_bar(stat="identity") + xlab("") + ylab("Proportion of lncRNA Triplets\n")+scale_fill_discrete(name = "lncRNA Modulator Type")
yy = yy+theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black"))
yy =yy+theme(text = element_text(size=20,color="black"),axis.text = element_text(size=20,color="black"))+scale_y_continuous(expand = c(0, 0))
yy

pdf("lncModulator_Type.pdf",width =10)
yy
dev.off()

```

# Figure 4c - lncRNA modulators and their tau scores 

```{r}

#Plot the tau score for shared modulators to see if being more ubiquitiously expressed correlates with functioning in more cancers as a modulator 

library(EnvStats)

tau_score = read.table("PanTARGET_AllGenes_TissueSpecificScore_040120.txt")
tau_score = tau_score[rownames(tau_score)%in%rownames(genes_anno),]
tau_score = tau_score[rownames(tau_score)%in%exp_cutoff_genes$V1,]
tau_score$gene_name = genes_anno[rownames(tau_score),4]
tau_score$gene_type = genes_anno[rownames(tau_score),3]
tau_score$gene_type_simple =as.character(tau_score$gene_type)
tau_score[grep("protein",tau_score$gene_type),5] = "protein"
tau_score[grep("lncRNA",tau_score$gene_type),5] = "lncRNA"
tau_score[grep("MSTRG_Non_coding_lncRNA",tau_score$gene_type),5] = "novel_lncRNA"
tau_score[grep("lnc-",tau_score$gene_name),5] = "novel_lncRNA"
tau_score = tau_score[grep("protein|lncRNA|novel_lncRNA",tau_score$gene_type_simple),]
tau_score$combo = paste(rownames(tau_score),tau_score$Tissue_name,sep="+")

pantarget_lncmod_lncRNAs = data.frame(table(as.character(pantarget_lncmod$lncRNA),as.character(pantarget_lncmod$cancer)))
pantarget_lncmod_lncRNAs = pantarget_lncmod_lncRNAs[pantarget_lncmod_lncRNAs$Freq!=0,]
#at this point there is information about how many cancer types per lncRNA are represented
pantarget_lncmod_lncRNAs = data.frame(table(as.character(pantarget_lncmod_lncRNAs$Var1)))
pantarget_lncmod_lncRNAs$tau_score = tau_score[as.character(pantarget_lncmod_lncRNAs$Var1),1]
#pancancer_lncmodulators = pantarget_lncmod_lncRNAs[pantarget_lncmod_lncRNAs$Freq>=4,1]
pantarget_lncmod_lncRNAs$Freq = factor(pantarget_lncmod_lncRNAs$Freq,c(1,2,3,4,5,6))


#combine the pan-cancer lncRNAs

#boxplot
zz =ggplot(pantarget_lncmod_lncRNAs,aes(pantarget_lncmod_lncRNAs$Freq,pantarget_lncmod_lncRNAs$tau_score))+geom_boxplot()+scale_fill_discrete(name="n")+xlab("\nNumber of Cancers with Shared lncRNA Modulator") + ylab("lncRNA Modulator Tau Score\n")+stat_n_text(size=5)
zz = zz+theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black"))
zz =zz+theme(text = element_text(size=20,color="black"),axis.text = element_text(size=20,color="black"))
zz

pdf("lncMod_TissueScore_SharedCancers_0420.pdf",width=8)
plot(zz)
dev.off()

```

# Figure 4d - lncRNA modulator target gene - gene set enrichment results and Supplementary Table 11

#MSIGDB Here

```{r}

#CHUNK9- Pathway analysis for lncMod target genes to assign pathway via guilt-by-association to lncRNA

#Msigdb analysis: - load the gene sets - you can download this from MSigDB
load("MSigDB_v5.0_human.rda")


#Functions created by Gonalazo to perform gene set enrichment using Fisher's Exact Test: 
fet=function(a,b,tot=NULL,alternative='t'){
## a =vector a
## b = vector b
## tot = background list
## alternative 't' = two-sided; 'g'=greater; 'l'=less
if(is.null(tot)) {
tot <- unique(c(a,b))
warning("No background list provided; a+b-(ab) used as background")
}
common <- length(intersect(intersect(a,b),tot))
#print(common)
na <- length(intersect(setdiff(a,intersect(a,b)),tot))
nb <- length(intersect(setdiff(b,intersect(a,b)),tot))
rest<- length(tot) - na - nb - common
#print(common)
#print(na)
#print(nb)
#print(rest)
res <- fisher.test(matrix(c(common,na,nb,rest),2,2),alternative=alternative)
return(list(p.value=res$p.value,oddsr=res$estimate,call=matrix(c(common,na,nb,rest),2,2)))
}

multi_fet <- function(set, collection, tot=NULL){
count =0
#set a vector containing gene names
#collection a list containing multiple pathway related gene sets
#background gene list (by default combines genes n the collection)
if(is.null(tot)) tot <- unique(unlist(collection))
res<-list()
for(i in names(collection)){
f <- fet(set,collection[[i]],tot)
#print(f)
res[[i]] <- c(f$p.value,f$oddsr, as.numeric(f$call))
}
ret <- data.frame(do.call(rbind,res))
fdr <- p.adjust(ret[,1],method="fdr")
ret <- cbind(fdr,ret)
colnames(ret) <- c("FDR","fet","OddsR","AB","BB","AA","tot")
return(ret[order(ret[,2]),])
}

#remove novel genes from triplet data

pantarget_lncmod = pantarget_lncmod[grep("MSTRG|lnc-",pantarget_lncmod$TF,invert=T),]
pantarget_lncmod = pantarget_lncmod[grep("MSTRG|lnc-",pantarget_lncmod$TargetGene,invert=T),]
#pantarget_lncmod = pantarget_lncmod[grep("MSTRG|lnc-",pantarget_lncmod$lncRNA,invert=T),]

#run the gene set enrichment per cancer individually:

lncmod_gsea = data.frame()

for (cancers in c("AML","TALL","BALL","RT","WT","NBL")){
  geneset = names(table(as.character(pantarget_lncmod[pantarget_lncmod$cancer==cancers,]$TargetGene)))
  geneset = genes_anno[geneset,4]
  geneset = geneset[grep("MSTRG|lnc-",geneset,invert=T)]
  background = genes_anno[rownames(pantarget[[cancers]]),4]
  background = background[grep("MSTRG|lnc-",background,invert=T)]
  #background = background[background%in%genes_anno[rownames(protein),4]]
  #KEGG or GO biological process
  #msigDBsymbol$c5.bp.v5.0.symbols.gmt
  #msigDBsymbol$c2.cp.kegg.v5.0.symbols.gmt
  #msigDBsymbol$h.all.v5.0.symbols.gmt
  collection=msigDBsymbol$h.all.v5.0.symbols.gmt
  #collection = msigDBsymbol$c5.bp.v5.0.symbols.gmt
  #collection = msigDBsymbol$c2.cp.kegg.v5.0.symbols.gmt
  #collection=NULL
  MSigDB_results = multi_fet(geneset,collection,background)
  MSigDB_results$cancer = cancers
  MSigDB_results$set = rownames(MSigDB_results)
  lncmod_gsea = rbind(lncmod_gsea,MSigDB_results)
}

MSigDB_results_sig = MSigDB_results[MSigDB_results$FDR<0.1,]
MSigDB_results_sig = MSigDB_results_sig[order(MSigDB_results_sig$fet,decreasing = T),]
MSigDB_results_sig$pathway = rownames(MSigDB_results_sig)
MSigDB_results_sig$pathway = factor(MSigDB_results_sig$pathway,levels=c(MSigDB_results_sig$pathway))

lncmod_gsea_sig = lncmod_gsea[lncmod_gsea$FDR<0.1,]
lncmod_gsea_sig$FDR2 = -log10(lncmod_gsea_sig$FDR)

lncmod_gsea$FDR2= -log10(lncmod_gsea$FDR)

#take the results for gene set enrichment above and make it into a table with columns as cancer and each row as a gene set 

lncmod_gsea_sig_plot = dcast(lncmod_gsea_sig,formula=set~cancer,value.var="FDR2")
#0720
lncmod_gsea_sig_plot = dcast(lncmod_gsea,formula=set~cancer,value.var="FDR2")
lncmod_gsea_sig_plot[is.na(lncmod_gsea_sig_plot)] = 0
rownames(lncmod_gsea_sig_plot) = lncmod_gsea_sig_plot$set
lncmod_gsea_sig_plot$set = NULL

#overall categories for the MsigDB gene sets 

hallcat = read.delim("MsigDB_Hallmark_Categories.txt",header=F)

rownames(hallcat) = hallcat$V1
hallcat = hallcat[order(hallcat$V2),]
hallcat = hallcat[!is.na(hallcat$V2),]
hallcat = hallcat[hallcat$V1%in%rownames(lncmod_gsea_sig_plot),]

MSigDB_results_sig$category = hallcat[rownames(MSigDB_results_sig),2]
lncmod_gsea_sig$category= hallcat[as.character(lncmod_gsea_sig$set),2]

lncmod_gsea_sig_plot = lncmod_gsea_sig_plot[as.character(hallcat$V1),]
lncmod_gsea_sig_plot = lncmod_gsea_sig_plot[,c("AML","BALL","TALL","RT","WT","NBL")]

rownames(lncmod_gsea_sig_plot) = gsub("HALLMARK_","",rownames(lncmod_gsea_sig_plot))

#rownames(lncmod_gsea_sig_plot) = tolower(rownames(lncmod_gsea_sig_plot))
#rownames(lncmod_gsea_sig_plot) = gsub("kegg_","",rownames(lncmod_gsea_sig_plot))

colors = colorRampPalette(c("white","orange","red"), space="rgb")(64)

yy =pheatmap(lncmod_gsea_sig_plot,show_rownames = T,show_colnames = T,cluster_cols = F,cluster_rows = F,color=colors,angle_col =0)
yy

pdf("GeneSet_Enrich_Heatmap_lncMod_TG_0720.pdf",width=5.5)
yy
dev.off()


#Supplementary Table 11 - with the pathway results 

#write.table(lncmod_gsea,"lncMod_MsigDB_Hallmark_results.txt",quote=F,row.names = F,col.names=F,sep="\t")

```

#Figure 4e-f - the top TFs for BLACE and the lncMod triplet heatmap for BLACE and XBP1

```{r}
library(ggplot2)
library(pheatmap)

path = read.delim("../Results/All_lncMod_Results_Anno.txt",header=T)

#BALL: BLACE-XBP1

plot_lncTF = data.frame(sort(table(path[path$lncRNA=="ENSG00000204960.5" & path$cancer=="BALL",]$TF_name),decreasing = T))
plot_lncTF = plot_lncTF[1:10,]

xx = ggplot(plot_lncTF,aes(Var1,Freq))+geom_bar(stat="identity")+ylab("# of Target Genes\n") + xlab("\nTranscription Factors")+ggtitle("B-ALL: BLACE")
xx = xx+theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black"))
xx =xx+theme(text = element_text(size=20,color="black"),axis.text = element_text(size=20,color="black"),axis.text.x = element_text(angle = 90, hjust = 1))
xx

pdf("BALL_BLACE_TFs.pdf")
xx
dev.off()

#heatmap for target genes for a specific lncRNA-TF.

cancer = BALL

temp = cancer[rownames(cancer)%in%as.character(pantarget_lncmod[pantarget_lncmod$lnc_TF=="BLACE_XBP1",]$TargetGene),]
temp = rbind(cancer["ENSG00000100219.12",],temp)
temp = rbind(cancer["ENSG00000204960.5",],temp)
name = genes_anno[rownames(temp),4]
temp = temp[,order(temp[1,])]

bottom = floor(length(cancer)*0.25)
top = (length(cancer)-floor(length(cancer)*0.25))+1
temp= temp[,c(1:bottom,top:length(cancer))]

temp[,1:bottom] = temp[,colnames(sort(temp[2,1:bottom]))]
temp[,(bottom+1):length(temp)] = temp[,colnames(sort(temp[2,(bottom+1):length(temp)]))]

triplet =temp

triplet = data.frame(t(scale(t(log2(triplet+1)),center=T,scale=T)))


temp_path = path[grep("BLACE_XBP1",path$triplet),]
rownames(temp_path) = temp_path$TargetGene

path$lnc_TF = paste(path$lncRNA_name,path$TF_name,sep="_")

row_anno = path[path$TargetGene%in%rownames(triplet[-2,]) & path$lnc_TF=="BLACE_XBP1",]$pathcat

row_anno = data.frame(c("","",as.character(row_anno)))

rownames(triplet)= name

rownames(row_anno)= name
colnames(row_anno) = "PathCat"
row_anno$extra = row_anno[,1]
row_anno = row_anno[order(row_anno$PathCat),]

triplet = triplet[rownames(row_anno),]

#triplet = triplet[c(2,3,1),]

colors = colorRampPalette(c("blue","white", "red"), space="rgb")(64)

aa=pheatmap(triplet,cluster_rows=F, show_colnames=F,show_rownames=T,cluster_cols=F, color=colors,fontsize = 16,gaps_col = 47,gaps_row = c(1,2),border_color = NA,annotation_row = row_anno)

pdf("BALL_BLACE_XBP1_Genes_Heatmap.pdf",height=9,width=15)
aa
dev.off()

```


# Supplementary Figure 5a - example of lncMod triplet 

```{r}
#WT: RMST-SOX2

cancer = WT

temp = cancer[rownames(cancer)%in%as.character(pantarget_lncmod[pantarget_lncmod$lnc_TF=="RMST_SOX2",]$TargetGene),]
temp = rbind(cancer["ENSG00000181449.2",],temp)
temp = rbind(cancer["ENSG00000255794.2",],temp)
name = genes_anno[rownames(temp),4]
temp = temp[,order(temp[1,])]

bottom = floor(length(cancer)*0.25)
top = (length(cancer)-floor(length(cancer)*0.25))+1
temp= temp[,c(1:bottom,top:length(cancer))]

temp[,1:bottom] = temp[,colnames(sort(temp[2,1:bottom]))]
temp[,(bottom+1):length(temp)] = temp[,colnames(sort(temp[2,(bottom+1):length(temp)]))]

triplet =temp

triplet = data.frame(t(scale(t(log2(triplet+1)),center=T,scale=T)))


rownames(triplet)= name

#triplet = triplet[c(2,3,1),]

colors = colorRampPalette(c("blue","white", "red"), space="rgb")(64)

aa=pheatmap(triplet,cluster_rows=F, show_colnames=F,show_rownames=T,cluster_cols=F, color=colors,fontsize = 16,gaps_col = bottom,gaps_row = c(1,2),border_color = NA)

pdf("WT_RMST_SOX2_Genes_Heatmap.pdf",height=9,width=15)
aa
dev.off()

```

#Supplementary Figure 5b - how many effects can a lncRNA modular have on target genes 

```{r}
#the lncMod type per lncRNA - We determine per unique lncRNA how many types of functions can it hvae depending on the triplet in question.

pantarget_lncmod$lnc_type_name = paste(pantarget_lncmod$lncRNA_name,pantarget_lncmod$lncMod_type,sep="+")

pantarget_lnctargets = data.frame(table(pantarget_lncmod$lnc_type_name,pantarget_lncmod$cancer))
pantarget_lnctargets = pantarget_lnctargets[pantarget_lnctargets$Freq!=0,]

pantarget_lnctargets = separate(data=pantarget_lnctargets,col="Var1",into=c("lncRNA","Type"),sep="\\+")

pantarget_lnctargets = data.frame(table(pantarget_lnctargets$lncRNA,pantarget_lnctargets$Var2))
pantarget_lnctargets = pantarget_lnctargets[pantarget_lnctargets$Freq!=0,]

#number of lncRNAs per type per cancer 
pantarget_lnctargets = data.frame(table(pantarget_lnctargets$Var2,pantarget_lnctargets$Freq))
table(pantarget_lnctargets$Freq)

pantarget_lnctargets$Var1 = factor(pantarget_lnctargets$Var1,levels=c("NBL","WT","BALL","TALL","AML","RT"))

yy = ggplot(pantarget_lnctargets,aes(pantarget_lnctargets$Var1,pantarget_lnctargets$Freq,fill=pantarget_lnctargets$Var2))+geom_bar(stat="identity")+xlab("")+ylab("Number of lncRNAs\n")+scale_fill_discrete(name = "Number of Modulator Types")
yy = yy+theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black"))
yy =yy+theme(text = element_text(size=20,color="black"),axis.text = element_text(size=20,color="black"))+scale_y_continuous(expand = c(0, 0))
yy

pdf("lncModulator_Type_PerLncRNA.pdf",width =10)
yy
dev.off()

```

#Supplementary Figure 5c - NBL H19 lncRNA modulatory gene set enrichment

```{r}
#NBL - H19

#Gene set enrichment for one lncRNA and its TF's target genes, i.e. for any one lncRNA in one cancer what are the pathways it appears to be enriched in based on its target genes. 

geneset = names(table(as.character(pantarget_lncmod[pantarget_lncmod$cancer=="NBL" & pantarget_lncmod$lncRNA_name=="H19",]$TargetGene)))
geneset = genes_anno[geneset,4]
geneset = geneset[grep("MSTRG|lnc-",geneset,invert=T)]

background = genes_anno[rownames(NBL),4]
background = background[grep("MSTRG|lnc-",background,invert=T)]

collection=msigDBsymbol$h.all.v5.0.symbols.gmt

#you will need to load this function from above: 

MSigDB_results = multi_fet(geneset,collection,background)
MSigDB_results$cancer = "NBL"
MSigDB_results$set = rownames(MSigDB_results)
MSigDB_results = MSigDB_results[1:5,]

MSigDB_results_sig = MSigDB_results[MSigDB_results$fet<0.05,]
MSigDB_results_sig = MSigDB_results_sig[order(MSigDB_results_sig$AB,decreasing = F),]
MSigDB_results_sig$pathway = rownames(MSigDB_results_sig)
MSigDB_results_sig$pathway = paste(MSigDB_results_sig$pathway,hallcat[as.character(MSigDB_results_sig$pathway),2],sep="_")
MSigDB_results_sig$pathway = gsub("HALLMARK_","",MSigDB_results_sig$pathway)
MSigDB_results_sig$pathway = factor(MSigDB_results_sig$pathway,levels=c(MSigDB_results_sig$pathway))


xx = ggplot(MSigDB_results_sig,aes(MSigDB_results_sig$pathway,MSigDB_results_sig$AB))+geom_bar(stat="identity",width=0.5)+coord_flip()+ylab("# of Target Genes\n") + xlab("\nHallmarks")+ggtitle("NBL: H19")
xx = xx+theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black"))
xx =xx+theme(text = element_text(size=20,color="black"),axis.text = element_text(size=20,color="black"),axis.text.x = element_text(angle = 0, vjust = 1))
xx
pdf("Hallmark_Pathways_NBL_H19_Enrichment.pdf",width=15)
xx
dev.off()

```

#Supplementary Figure 5d - NBL H19 lncRNA modulatory gene set enrichment

```{r}
#AML - HOTAIRM1

#Gene set enrichment for one lncRNA and its TF's target genes, i.e. for any one lncRNA in one cancer what are the pathways it appears to be enriched in based on its target genes. 

geneset = names(table(as.character(pantarget_lncmod[pantarget_lncmod$cancer=="AML" & pantarget_lncmod$lncRNA_name=="HOTAIRM1",]$TargetGene)))
geneset = genes_anno[geneset,4]
geneset = geneset[grep("MSTRG|lnc-",geneset,invert=T)]

background = genes_anno[rownames(AML),4]
background = background[grep("MSTRG|lnc-",background,invert=T)]

MSigDB_results = multi_fet(geneset,collection,background)
MSigDB_results$cancer = "AML"
MSigDB_results$set = rownames(MSigDB_results)

#you will need to load this function from above: 

MSigDB_results_sig = MSigDB_results[MSigDB_results$fet<0.05,]
MSigDB_results_sig = MSigDB_results_sig[order(MSigDB_results_sig$AB,decreasing = F),]
MSigDB_results_sig$pathway = rownames(MSigDB_results_sig)
MSigDB_results_sig$pathway = paste(MSigDB_results_sig$pathway,hallcat[as.character(MSigDB_results_sig$pathway),2],sep="_")
MSigDB_results_sig$pathway = gsub("HALLMARK_","",MSigDB_results_sig$pathway)
MSigDB_results_sig$pathway = factor(MSigDB_results_sig$pathway,levels=c(MSigDB_results_sig$pathway))

xx = ggplot(MSigDB_results_sig,aes(MSigDB_results_sig$pathway,MSigDB_results_sig$AB))+geom_bar(stat="identity",width=0.5)+coord_flip()+ylab("# of Target Genes\n") + xlab("\nHallmarks")+ggtitle("AML: HOTAIRM1")
xx = xx+theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black"))
xx =xx+theme(text = element_text(size=20,color="black"),axis.text = element_text(size=20,color="black"),axis.text.x = element_text(angle = 0, vjust = 1))
xx

pdf("Hallmark_Pathways_AML_HOTAIRM1_Enrichment.pdf",width=10)
xx
dev.off()

```

#Supplementary Figure 5e - Heatmap and TF bar graph - GAS5+E2F4

```{r}
#T-ALL: GAS5

path = read.delim("All_lncMod_Results_Anno.txt",header=T)

plot_lncTF = data.frame(sort(table(path[path$lncRNA=="ENSG00000234741.3" & path$cancer=="TALL",]$TF_name),decreasing = T))
plot_lncTF = plot_lncTF[1:10,]

xx = ggplot(plot_lncTF,aes(Var1,Freq))+geom_bar(stat="identity")+ylab("# of Target Genes\n") + xlab("\nTranscription Factors")+ggtitle("T-ALL: SNHG1")
xx = xx+theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black"))
xx =xx+theme(text = element_text(size=20,color="black"),axis.text = element_text(size=20,color="black"),axis.text.x = element_text(angle = 90, hjust = 1))
xx

pdf("TALL_GAS5_TFs.pdf")
xx
dev.off()

#heatmap for proliferation genes for a specific lncRNA-TF. Need to have run chunk9/10.

cancer = TALL

temp = pantarget_lncmod[pantarget_lncmod$cancer=="TALL" & pantarget_lncmod$lnc_TF=="GAS5_E2F4",]
temp = temp[order(temp$Bottom_cor),]
temp = cancer[as.character(temp$TargetGene),]

temp = rbind(cancer["ENSG00000205250.4",],temp)
temp = rbind(cancer["ENSG00000234741.3",],temp)
name = genes_anno[rownames(temp),4]
temp = temp[,order(temp[1,])]

bottom = floor(length(cancer)*0.25)
top = (length(cancer)-floor(length(cancer)*0.25))+1
temp= temp[,c(1:bottom,top:length(cancer))]

temp[,1:bottom] = temp[,colnames(sort(temp[2,1:bottom]))]
temp[,(bottom+1):length(temp)] = temp[,colnames(sort(temp[2,(bottom+1):length(temp)]))]

triplet =temp

triplet = data.frame(t(scale(t(log2(triplet+1)),center=T,scale=T)))

rownames(triplet)= name

#triplet = triplet[c(2,3,1),]

colors = colorRampPalette(c("blue","white", "red"), space="rgb")(64)

aa=pheatmap(triplet,cluster_rows=F, show_colnames=F,show_rownames=T,cluster_cols=F, color=colors,fontsize = 16,gaps_col = bottom,gaps_row=c(1,2),border_color = NA)
aa

pdf("TALL_GAS5_E2F4_Genes_Heatmap.pdf",height=9,width=15)
aa
dev.off()

```

#Supplementary Figure 5e - Heatmap and TF bar graph - SNHG1+TP53

```{r}
#T-ALL: SNHG1_TP53

plot_lncTF = data.frame(sort(table(path[path$lncRNA=="ENSG00000255717.2" & path$cancer=="TALL",]$TF_name),decreasing = T))
plot_lncTF = plot_lncTF[1:10,]

xx = ggplot(plot_lncTF,aes(Var1,Freq))+geom_bar(stat="identity")+ylab("# of Target Genes\n") + xlab("\nTranscription Factors")+ggtitle("T-ALL: SNHG1")
xx = xx+theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black"))
xx =xx+theme(text = element_text(size=20,color="black"),axis.text = element_text(size=20,color="black"),axis.text.x = element_text(angle = 90, hjust = 1))
xx

pdf("TALL_SNHG1_TFs.pdf")
xx
dev.off()


#heatmap for proliferation genes for a specific lncRNA-TF. Need to have run chunk9/10.

cancer = TALL

temp = pantarget_lncmod[pantarget_lncmod$cancer=="TALL" & pantarget_lncmod$lnc_TF=="SNHG1_TP53",]
temp = temp[order(temp$Bottom_cor),]
temp = cancer[as.character(temp$TargetGene),]

temp = rbind(cancer["ENSG00000141510.11",],temp)
temp = rbind(cancer["ENSG00000255717.2",],temp)
name = genes_anno[rownames(temp),4]
temp = temp[,order(temp[1,])]

bottom = floor(length(cancer)*0.25)
top = (length(cancer)-floor(length(cancer)*0.25))+1
temp= temp[,c(1:bottom,top:length(cancer))]

temp[,1:bottom] = temp[,colnames(sort(temp[2,1:bottom]))]
temp[,(bottom+1):length(temp)] = temp[,colnames(sort(temp[2,(bottom+1):length(temp)]))]

triplet =temp

triplet = data.frame(t(scale(t(log2(triplet+1)),center=T,scale=T)))

rownames(triplet)= name

#triplet = triplet[c(2,3,1),]

colors = colorRampPalette(c("blue","white", "red"), space="rgb")(64)

aa=pheatmap(triplet,cluster_rows=F, show_colnames=F,show_rownames=T,cluster_cols=F, color=colors,fontsize = 16,gaps_col = bottom,gaps_row=c(1,2),border_color = NA)
aa

pdf("TALL_SNHG1_TP53_Genes_Heatmap.pdf",height=9,width=15)
aa
dev.off()

```

#Supplementary Table 12

```{r}

library(dplyr)
library(reshape2)
#create table of lnc-TF associations for supplementary table

load("PanTarget_lncMod_Results_0420.rda")

pantarget_lncmod$lnc_TF = gsub("\\+","_",pantarget_lncmod$lnc_TF)
pantarget_lncmod$lnc_TG = gsub("\\+","_",pantarget_lncmod$lnc_TG)

lncTF = data.frame(table(pantarget_lncmod$cancer,pantarget_lncmod$lncRNA_name,pantarget_lncmod$TF_name))
lncTF = lncTF[lncTF$Freq!=0,]

lncTF = lncTF[order(lncTF$Freq,decreasing=T),]
lncTF = lncTF %>% group_by(Var1,Var2) %>% mutate(rank=row_number())
lncTF$cancer_lnc = paste(lncTF$Var1,lncTF$Var2,sep=": ")
lncTF$TF_Freq = paste(lncTF$Var3,lncTF$Freq,sep=": ")
lncTF = dcast(lncTF,cancer_lnc~rank,value.var = "TF_Freq")
lncTF = lncTF[,c(1:11)]

write.table(lncTF,"Supp_Table12_lncTF_lncMod.txt",quote=F,row.names=F,sep="\t")

```
